### 2015-05-01 paper/main.tex
updated paper

\pagestyle{myheadings}
%% ** EDIT HERE **
\usepackage{enumerate}
\usepackage{subfig}
\usepackage{multirow} 
\usepackage{url}
\usepackage{xr} %for cross-referencing
%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\newtheorem{algorithm}{Algorithm}
\newtheorem{proposition}{Proposition}
\newtheorem{restateproposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{result}{Result}
\newtheorem{note}{Note}
\newtheorem{definition}{Definition}

@@

Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
EB approaches to this problem have been quite extensively studied by several authors, 
especially B.~Efron and co-authors 
\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
see also \cite{kendziorski2003parametric,newton2004detecting, 
datta2005empirical,muralidharan2010empirical} for example.
So what is the ``New Deal" here? We introduce two simple new ideas that can substantially improve
inference. The first idea is to {\it assume that the distribution of the non-zero effects is unimodal}, and to perform
non-parametric inference under this assumption. This idea has the twin benefits of both simplifying and
improving inference, at least when the assumption is correct. And even when the assumption is incorrect, we show
that the approach can be relatively robust to this. The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
 to summarize each measurement. This idea allows any differences in the precision of different measurements to be taken into account,
 and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements, resulting in inflated FDR values.

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle Bayesian approaches, including EB, can naturally unify testing and estimation into a single framework -- testing is
simply estimation but with a point mass of some weight on the null hypothesis that an effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice, large-scale studies that assess many effects almost invariably focus on testing significance, and
controlling the FDR, rather than on estimating the sizes of the effect. 
The two new ideas highlighted above 
help facilitate estimation as well as testing, and by providing simple software to implement our methods
we hope to promote and accelerate the shift in this direction.

Here we consider one setting in which FDR-based methods are widely used:
an experimenter has measured a large numbers of ``effects", each with an associated
measurement precision (standard error), 
and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
This setting arises commonly in many areas of statistics, and particularly in genomics applications.
For example, in genomics
a common goal is to identify ``differentially-expressed genes" -- that is, genes
whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
mean expression level between the two conditions, often referred to as the ``log-fold change". 
%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.

@@

Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
Here we present new tools that bring two important ideas to bear on the problem.
First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
a substantial increase in the number of discoveries at a given FDR).

Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests, 
provides a general tool for shrinkage estimation. Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
where shrinkage of wavelet coefficients is used to smooth signals. 
However, in contrast to FDR methods, shrinkage methods are not so widely used in practice, at least in the kinds
of genomics settings I am familiar with. One possible reason for this is that most work on shrinkage-based estimation has focussed
on point estimates, rather than on interval estimates which are often of more direct practical relevance. Another contributing
factor may be the paucity of general-purpose software packages for this kind of shrinkage estimation.
% (There are, perhaps, more packages for shrinkage estimation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
Our work here addresses both these issues, providing a general-purpose software tool that uses shrinkage ideas to improve both
point and interval estimates.

By ``adaptive" shrinkage we 
have two properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with high standard error will undergo more shrinkage than measurements with low standard error.  
Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements to improve inference for each individual effect. By improved inference, we mean both
improved average accuracy of point estimates, which is
the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty, such as credible intervals.

The potential for shrinkage-based estimation to
address the multiple comparisons problems tackled by FDR-based methods
has been highlighted
previously, including \cite{greenland1991empirical},
\cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984] 
Furthermore,

The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\bhat=(\bhat_1,\dots, \bhat_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).

Given that shrinkage estimation is widely recognized as a powerful tool, there are

The only package we have found that provides anything similar
to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of \ashr are that it
i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).

As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
generic methods that work directly with two numbers for each 
measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.

@@

Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$). In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
 We assume that we have obtained data $D=(D_1,\dots,D_J)$ that provide independent estimates $\bhat_1,\dots,\bhat_J$ of these effects,
 with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds, and $t_j$ denote
 the $t$ statistic testing this null:
 \begin{gather}
 H_j: \beta_j = 0 \\  
 t_j :=\bhat_j/\shat_j. 
 \end{gather}
 The $t$ statistics, of course, have a $t$ distribution under the null:
 \begin{equation} \label{eqn:tstat}
 t_j | H_j \sim T_\nu
 \end{equation}
 where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. 
 Let $p_j$ denote the $p$ value obtained by comparing $t_j$ with its null distribution,
 \begin{equation}
 p_j := \Pr(|T_\nu| > |t_j| ).
 \end{equation}
Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values;
 these methods can be applied by translating each $t_j$ into a corresponding
$z$ score, which we denote $z_j$.

@@

Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
\begin{equation} \label{eqn:FDR}
\FDR(\Gamma) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
\end{equation}
taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
the expectation of $\FDR(\Gamma)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
%\begin{equation} \label{eqn:FDR}
%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
%\end{equation}
\begin{equation} \label{eqn:fdr}
\fdr_j := \Pr(H_j | D).
\end{equation}
The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
\begin{equation} 
\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
\end{equation}

The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
\begin{equation}
q_j := \FDR(\{k:\fdr_k \leq \fdr_j\}).
\end{equation}
For existing EB methods (but not for our method), typically $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
\begin{equation}
q_j = \FDR(\{k:p_k \leq p_j\})
\end{equation}
which is close to the definition in \cite{storey.03}.

One nice feature of FDR procedures is that Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
for example \cite{storey.03}, particularly his Theorem 1.

\subsection{Existing methods}

\subsubsection{qvalue}

@@

\subsubsection{EB approaches: \locfdr and \mixfdr}

The fundamental
 idea of EB procedures is similar to that of \qvalue: to attempt to partition the observations into null and alternative components.
Here we highlight two EB methods that
 seem most relevant to our work here and which had good software implementations at the time of writing. 
 The first is due to \cite{efron2008microarrays}, implemented in the R package \locfdr.
 This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
 %Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
 %one corresponding to the null tests, and the other corresponding to the alternative tests. 
 In the simplest case
 \locfdr  makes assumptions analogous to the two key assumptions of \qvalue: i) the null $z$ scores have a $N(0,1)$ distribution,
 and ii) all the $z$ scores near 0 are from the null, which Efron
calls the ``Zero Assumption" (ZA).  
The \locfdr package first estimates the overall distribution of $z$ scores using a non-parametric regression approach, and then, using the two assumptions,  ``subtracts out" the
null component, leaving the remainder to be the alternative component. In this way \locfdr, like \qvalue, avoids explicit modelling of the alternative component.
 (An important focus of Efron's work, also implemented in \locfdr, 
 is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.)

@@

 Here we introduce an EB method, which, for reasons discussed below, we call ``adaptive shrinkage" or ``ash".
 The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates, shrinkage credible intervals,
 and estimates of the local false sign rate, defined below. It is implemented in the R package \ashr available at \url{http://github.com/stephens999/ashr}.

  Rather than directly modelling the distributions of the $p$ values, or $z$ scores, our starting point is to model the distribution of the effects $\beta_j$.
 Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
  \begin{equation} \label{eqn:beta}
 \beta_j \sim g(\cdot).
 \end{equation}
 A key assumption, which distinguishes our approach from previous work, is that $g$ is unimodal. We assume here that the mode is at 0, and allow
 that $g$ can have a point mass at 0, corresponding to the ``null" effects.
(An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data.)

 \def\dt{\text{dt}}

To perform likelihood-based inference for $g,\beta$, we assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
 $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
  \begin{equation} \label{eqn:lik}
 \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
 \end{equation}
 where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e. the density of $\mu+\sigma T_\nu$). In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
  \begin{equation} \label{eqn:normlik}
 \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
 \end{equation}
%which, in light of (\ref{eqn:betahat}), is equivalent to
 % \begin{equation} \label{eqn:lik2}
% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
 %\end{equation}

 \subsubsection*{Discussion on Assumptions}

Although the unimodal assumption (UA) will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
For example, in contexts where interest focusses on which $\beta_j$ differ from zero, which suggests that ``$\beta_j=0$" is a plausible null hypothesis,  
it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible, and that the distribution of the effects will be unimodal about 0. 
Alternatively, we can motivate  the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
such shrinkage is desirable from several standpoints for improving estimation accuracy. 
Note that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero). It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.

%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
Note that, although the UA appears novel in this context,
almost all analogous work in large-scale regression make the UA for the regression coefficents - 
common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{}.
One of our aims here is to emphasize the considerable practical benefits of making the UA in the FDR context.

The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$. The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$. Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$. If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$. See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.

>>>>

\pagestyle{myheadings}
%% ** EDIT HERE **
\usepackage{enumerate}
\usepackage{multirow} 
\usepackage{url}
\usepackage{xr} %for cross-referencing
%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\newtheorem{algorithm}{Algorithm}
\newtheorem{proposition}{Proposition}
\newtheorem{restateproposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{result}{Result}
\newtheorem{note}{Note}
\newtheorem{definition}{Definition}

@@

Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
have been extensively studied by several authors, 
especially B.~Efron and co-authors 
\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
see also \cite{kendziorski2003parametric,newton2004detecting, 
datta2005empirical,muralidharan2010empirical} for example.
So what is the ``New Deal" here? We introduce two simple new ideas that can substantially improve
inference. The first idea is to {\it assume that the distribution of the non-zero effects is unimodal}, and to perform
non-parametric inference under this assumption. This idea has the twin benefits of simplifying and improving inference,
at least when the unimodal assumption is correct.  And even when the assumption is incorrect, the resulting errors in inference are not too egregious. 
The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
 to summarize each measurement. This idea allows any differences in the precision of different measurements to be taken into account,
 and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements and inflate the FDR.

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle Bayesian approaches, including EB, can naturally unify testing and estimation into a single framework -- testing is
simply estimation but with a point mass of some weight on the null hypothesis that an effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice, large-scale studies that assess many effects almost invariably focus on testing significance, and
controlling the FDR, rather than on estimating the sizes of the effect. 
The two new ideas highlighted above facilitate estimation as well as testing, and by providing easy-to-use software implementing our methods
we hope to promote and accelerate the shift in this direction.

The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron.morris.73.jasa}.
This close relationship becomes especially intimate here due to our assumption that the distribution of the underlying effects is
unimodal, since this ensures that estimates are shrunk towards a common point -- the mode of the distribution.
%Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
%if effects are distributed uni-modally about zero then it is natural to shrink estimates towards zero. 
% are essentially two sides of the same coin, one related to testing and the other to estimation. 
Thus, although we focus here primarily on FDR-related issues, another important contribution of our work is to provide {\it generic} and {\it adaptive} shrinkage estimation procedures.
By generic, we mean that these methods can be applied in any setting where a series of effect estimates and corresponding standard errors are available.
By adaptive, we 
have two properties in mind. First, the appropriate amount of shrinkage is determined from the data: when the data indicate that large effects are relatively common then
shrinkage of large observed effects is less than when the data indicate that large effects are rare.
Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors will undergo more shrinkage than measurements with small standard errors.  
Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
to existing EB shrinkage methods for that context, as is explored more fully elsewhere (Xing and Stephens).

Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.

%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.

@@

%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
%Here we present new tools that bring two important ideas to bear on the problem.
%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
%a substantial increase in the number of discoveries at a given FDR).

%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
%generic methods that work directly with two numbers for each 
%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.

@@

% Here we consider one setting in which FDR-based methods are widely used:
%an experimenter has measured a large numbers of ``effects", each with an associated
%measurement precision (standard error), 
%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
%For example, in genomics
%a common goal is to identify ``differentially-expressed genes" -- that is, genes
%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
%mean expression level between the two conditions, often referred to as the ``log-fold change".

Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$). For example, in a canonical genomics application
$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
In some contexts interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts interest may focus on estimating their values; the methods described here tackle both these problems.
 We assume that we have obtained data $D$ that consist of estimates $\bhat_1,\dots,\bhat_J$ of these effects,
 with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds ($H_j:\beta_j=0$), and $t_j$ denote
 the $t$ statistic testing this null:
 \begin{equation}  
 t_j :=\bhat_j/\shat_j. 
 \end{equation}
 The $t$ statistics, of course, have a $t$ distribution under the null:
 \begin{equation} \label{eqn:tstat}
 t_j | H_j \sim T_\nu
 \end{equation}
 where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. 
 Let $p_j$ denote the $p$ value obtained by comparing $t_j$ with its null distribution,
 \begin{equation}
 p_j := \Pr(|T_\nu| > |t_j| ).
 \end{equation}
Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values;
 these methods can be applied by translating each $t_j$ into a corresponding
$z$ score, which we denote $z_j$.

@@

Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
\begin{equation} \label{eqn:FDR}
\FDR(\Gamma) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
\end{equation}
taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
the expectation of $\FDR(\Gamma)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
%\begin{equation} \label{eqn:FDR}
%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
%\end{equation}
\begin{equation} \label{eqn:fdr}
\fdr_j := \Pr(H_j | D).
\end{equation}
The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
\begin{equation} 
\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
\end{equation}
Although our approach here is (Empirical) Bayesian, one nice feature of FDR procedures is that 
Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
for example \cite{storey.03}, particularly his Theorem 1.

The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
\begin{equation}
q_j := \FDR(\{k:\fdr_k \leq \fdr_j\}).
\end{equation}
In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
\begin{equation}
q_j = \FDR(\{k:p_k \leq p_j\})
\end{equation}
which is close to the definition in \cite{storey.03}.

\subsection*{Existing methods for FDR analysis}

\subsubsection*{qvalue}

@@

\subsubsection*{EB approaches: \locfdr and \mixfdr}

The fundamental
 idea of EB procedures is similar to that of \qvalue: partition the observations into null and alternative components.
Here we highlight two EB methods that
 seem most relevant to our work here and which had good software implementations at the time of writing. 
 The first is due to \cite{efron2008microarrays}, implemented in the R package \locfdr.
 This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
 %Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
 %one corresponding to the null tests, and the other corresponding to the alternative tests. 
 In the simplest case
 \locfdr  makes assumptions analogous to the two key assumptions of \qvalue: i) the null $z$ scores have a $N(0,1)$ distribution,
 and ii) all the $z$ scores near 0 are from the null, which Efron
calls the ``Zero Assumption" (ZA).  
The \locfdr package first estimates the overall distribution of $z$ scores using a non-parametric regression approach, and then, using the two assumptions,  ``subtracts out" the
null component, leaving the remainder to be the alternative component. In this way \locfdr, like \qvalue, avoids explicit modelling of the alternative component.
 (An important focus of Efron's work, also implemented in \locfdr, 
 is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.)

@@

 Here we introduce an EB method, which, for reasons discussed above, we call ``adaptive shrinkage" or ``ash".
 The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates and credible intervals for $\beta_j$,
 and estimates of the local false sign rate, defined below.

We begin by presenting the simplest version of our model, before embellishing it.
Instead of attempting to partition the (observed) $z$ scores or $p$ values into null and alternative components,
we use a model that partitions the (unobserved) effects $\beta_j$  into null and alternative components:
\begin{equation} \label{eqn:betaj}
\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)
\end{equation}
where $\delta_0$ denotes a point mass at 0, and $g_1$ denotes the (unknown) distribution of the non-zero $\beta_j$.
Thus $\pi_0$ denotes the proportion of effects that are null.

At this point we introduce a key assumption, which distinguishes our approach from previous work: we assume that $g_1$ is unimodal,
with a mode at 0. One simple way to achieve this ``Unimodal Assumption" (UA)  is
to assume that $g_1$ is a mixture of zero-mean normal distributions
\begin{equation} \label{eqn:g1}
g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
\end{equation}
where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
 The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We discuss 
 more flexible alternatives based on using mixtures of uniform distributions, and details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$), as well as the UA more generally, below.

Since the $\beta_j$ are unobserved, we need a model to connect the $\beta_j$ to observations. Here we
treat the estimates $\bhat_j$ as the observed data, and assume that they are noisy measurements of the true $\beta_j$ with given standard error
$\shat_j$:
\begin{equation} \label{eqn:bhat}
\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
\end{equation}
This approach, which follows \cite{wakefield:2009}, 
should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.

Combining (\ref{eqn:betaj})-(\ref{eqn:bhat}), and assuming independence of the observations and effects for different $j$ (given $\pi_0,w_1,\dots,w_k$), yields
a hierarchical model for $\beta,\bhat$ given $\shat$, with unknown hyper-parameters $(\pi_0,w_1,\dots,w_k)$. To simplify notation we define $\pi_k := (1-\pi_0)w_k$ 
so that the unknown hyper-parameters can be written as a vector $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1.
The usual empirical Bayes approach to fitting this model would involve two steps:
\begin{enumerate}
\item Estimate the hyperparameters $\pi$ by maximum likelihood, yielding a maximum likelihood estimate $\hat{\pi}$.
\item Compute quantities of interest (e.g. local fdr values, $q$ values, credible intervals etc.) from the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$.
\end{enumerate}
We make one modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
likelihood, where the penalty (detailed below) encourages $\hat\pi_0$ to be as big as possible (within the constrains of the model and observed data). We use
this penalty here because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR. 
Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{EM} to maximize the penalized likelihood, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

\subsubsection*{Embellishments}

\subsubsection*{More flexible prior component distributions}

\subsubsection*{Replace normal likelihood with $t$ likelihood}

\subsubsection*{}
